Efficiency

### 2018-01-08 DraftExec.Rnw
check if you like this style...

@@

In order to achieve a reduction in cavitation, however, care must be taken in the design of a decelerating duct.  Since the negative thrust produced by the duct must necessitate additional work being done by the impeller, there can be circumstances where the additional loading of the impeller out-weighs the impacts of elevating the pressure around it. \cite[24-25]{oosterveld1970} provides a more thorough derivation of the circumstances in which this can be reliably achieved, which generally include larger blade area ratios or with more rows of rotors for higher loading, or circumstances where the loading of the propeller is quite low.  This creates an inevitable and necessary tension between optimisation of a duct for propulsive efficiency and reduction of cavitation, which will be returned to in subsequent section.

@@

\subsubsection{Waterjets}
The use of the word waterjet, as opposed to pumpjet, has a tendency signify that the jet expelled from the shroud or pump is situated above the water, which is desirable in very fast-moving surface vessels which attempt to minimise their contact with the waters surface, including high-speed planing hull vessels, including pleasure craft.  The physics of the two systems, however, are closely related, with the very noticeable exception of the absence of flow considerations for water on the exterior of the duct considerations for waterjets.  However, the physics relating to the impeller and stator blades or (turbomachinery) designed to add energy to the flow are generally extremely closely linked, though a larger variety of diffeent pump types can be chosen for waterjets which might differ more significantly from the natural evolution of a propeller to an axial-flow pump. Many of the essential equations which govern the efficiencies and cavitation performance of pumpjets and waterjets are of the same or extremely similar form, as can be seen in detailed discussions given to both, by similar authors (\cite{wislicenus1973}, \cite{henderson1964}).  There are, however, plenty of instances in recent literature where this distinction regarding the ejection of the jet above and below the waterline is not maintained, and waterjets and pumpjets seem to be used interchangeably \parencite{abdel2010}, or systems are described as 'submerged waterjets' \parencite{buckingham2008submarine}.

@@

There are other some fundamental limitations on the peak efficiency that can be achieved by jets based on simple calculations of ideal efficiency from momentum theory, which are discussed in Section \ref{tradeoffs}.  They also provide enduring, reliable limitations on the extent to which ingenuity or design changes might be able to improve the efficiency of jets at low speed. But the primary reason why jet efficiencies decline towards zero at low speed is to do with flow separation \parencite{fujisawa1995}, which is what occurs when the assumptions inherent in momentum theory's analysis of an 'ideal' actuator disc start to break down.

@@

Flow separation also tends to lead to unsteady flows, with vortices periodically being shed into the main stream as they are speed up or change shape, and then ingested by the impeller and stator blades.  As can be seen by the diagram in Figure \ref{fig:JetLowFlow.png}, regions abound which are adjacent to or behind flow blockages or reversing flow, where considerable turbulence and unsteady forces would exist.  These effects also have acoustic consequences, which will be discussed in Section \ref{acousticconsiderations}.

>>>>

\usepackage{nameref}

@@

In order to achieve a reduction in cavitation, however, care must be taken in the design of a decelerating duct.  Since the negative thrust produced by the duct must necessitate additional work being done by the impeller, there can be circumstances where the additional loading of the impeller out-weighs the impacts of elevating the pressure around it. \cite[24-25]{oosterveld1970} provides a more thorough derivation of the circumstances in which this can be reliably achieved, which generally include larger blade area ratios or with more rows of rotors for higher loading, or circumstances where the loading of the propeller is quite low.  This creates an inevitable and necessary tension between optimisation of a duct for propulsive efficiency and reduction of cavitation, which will be returned to in Section \ref{waterjets} (\nameref{waterjets}).

@@

\subsubsection{Waterjets} \label{waterjets}

The use of the word waterjet, as opposed to pumpjet, has a tendency signify that the jet expelled from the shroud or pump is situated above the water, which is desirable in very fast-moving surface vessels which attempt to minimise their contact with the waters surface, including high-speed planing hull vessels, including pleasure craft.  The physics of the two systems, however, are closely related, with the very noticeable exception of the absence of flow considerations for water on the exterior of the duct considerations for waterjets.  However, the physics relating to the impeller and stator blades or (turbomachinery) designed to add energy to the flow are generally extremely closely linked, though a larger variety of diffeent pump types can be chosen for waterjets which might differ more significantly from the natural evolution of a propeller to an axial-flow pump. Many of the essential equations which govern the efficiencies and cavitation performance of pumpjets and waterjets are of the same or extremely similar form, as can be seen in detailed discussions given to both, by similar authors (\cite{wislicenus1973}, \cite{henderson1964}).  There are, however, plenty of instances in recent literature where this distinction regarding the ejection of the jet above and below the waterline is not maintained, and waterjets and pumpjets seem to be used interchangeably \parencite{abdel2010}, or systems are described as 'submerged waterjets' \parencite{buckingham2008submarine}.

@@

There are other some fundamental limitations on the peak efficiency that can be achieved by jets based on simple calculations of ideal efficiency from momentum theory, which are discussed in Section \ref{tradeoffs} (\nameref{tradeoffs}).  They also provide enduring, reliable limitations on the extent to which ingenuity or design changes might be able to improve the efficiency of jets at low speed. But the primary reason why jet efficiencies decline towards zero at low speed is to do with flow separation \parencite{fujisawa1995}, which is what occurs when the assumptions inherent in momentum theory's analysis of an 'ideal' actuator disc start to break down.

@@

Flow separation also tends to lead to unsteady flows, with vortices periodically being shed into the main stream as they are speed up or change shape, and then ingested by the impeller and stator blades.  As can be seen by the diagram in Figure \ref{fig:JetLowFlow.png}, regions abound which are adjacent to or behind flow blockages or reversing flow, where considerable turbulence and unsteady forces would exist.  These effects also have acoustic consequences, which will be discussed in Section \ref{acousticconsiderations} (\nameref{acousticconsiderations}).
